Community-Centric Innovation

### 2014-02-21 paper.tex
other projects

\subsection{Networked Cryptocurrencies}

Another area that has witnessed the potential for community-centric innovation is that of networked cryptocurrency.
The combination of cryptography, free libre software and a data network can result in networked cryptocurrencies.
In this technologies, a cryptographically secure distributed ledger is maintained by the network.
The network is composed of nodes that interact in a peer-to-peer fashion following a common set of rules or protocol to guarantee the integrity of the cryptocurrency.
The network not only distributes information about the ongoing transactions, it also provides hashing power that is used as a proof of work.
The distributed ledger is protected by the proof of work.
In a sense, the nodes vote with their proof-of-work and the majority decides which transactions are included in the public distributed ledger.

Cryptocurrency has some advantages when compared to traditional fiat currencies.
Some of the advantages of cryptocurrencies are related to their peer-to-peer nature.
One of them is the lack of a central authority that can make arbitrary decisions and create money at no cost.
The emission of money is controlled by the protocol and governed by publicly known rules.
This means that the availability of money at any point in the future is known by all the participants.
The emitted money is used to reward the nodes that contribute with hashing power to the security of the network.
Hashing power requires hardware and electricity, which means that is no longer possible to create money at zero cost.
In fact, it is a risky endeavour as the cost of hashing power can exceed the value of the generated money.

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
There is no risks of blockades of political interference.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency to bypass totalitarian governments blockades or lacking payment networks.

@@

As cryptocurrencies are based on free libre software, anyone can create a new currency.
If a currency is used by a single individual is not very useful.
The potential of networked cryptocurrencies evidences when they are used by communities.
The communities can create or agree in a currency for their internal exchanges.
As a result, the currency is impregnated by the values and ethical standards of the community.

The people can choose which currency to use and which community to join.
In this choice, there is an implicit commitment to support certain values and principles.
By joining a particular community and currency, the user strengthens that particular community by offering productive force and services.
The currency is a commons resource used by the whole community.
The value of currency increases as the number of users and transactions increases.
Therefore, the members of a community have an incentive to devote some currency resources to attract new members to the community.

>>>>

\section{Networked Cryptocurrencies}

Another area that has witnessed the potential for community-centric innovation is that of networked cryptocurrency.
The combination of cryptography, free libre software and a data network make it possible the appearance of networked cryptocurrencies.
In these technologies, a cryptographically secure distributed ledger is maintained by the network.
The network is composed of nodes that interact in a peer-to-peer fashion following a common set of rules or protocol to guarantee the integrity of the cryptocurrency.
The network not only distributes information about the ongoing transactions, it also provides hashing power that is used as a proof of work.
The distributed ledger is protected by the proof of work.
In a sense, the nodes vote with their proof-of-work and the majority decides which transactions are included in the public distributed ledger.

Cryptocurrency has some advantages when compared to traditional fiat currencies.
Some of the advantages of cryptocurrencies are related to their peer-to-peer nature.
One of them is the lack of a central authority that can make arbitrary decisions and create money at no cost.
In cryptocurrencies, the emission of money is controlled by the protocol and governed by publicly known rules.
This means that the availability of money at any point in the future is known by all the participants.
The emitted money is used to reward the nodes that contribute with hashing power to the security of the network.
Hashing power requires hardware and electricity, which means that is no longer possible to create money at zero cost.
In fact, the money generation business (popularly known as \emph{mining}) is a risky endeavour as the cost of hashing power can exceed the value of the generated money.

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
The risks of blockades or political interference are greatly mitigated.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency to bypass totalitarian governments blockades or lacking payment networks.

@@

As cryptocurrencies are based on free libre software, anyone can create a new currency.
If a currency is used by a single individual is not very useful.
The potential of networked cryptocurrencies materializes when they are used by communities.
The communities can create or agree in a currency for their internal exchanges.
As a result, the currency is impregnated by the values and ethical standards of the community.

The people can choose which currency to use and which community to join.
In this choice, there is an implicit commitment to support certain values and principles.
By joining a particular community and currency, the user strengthens that particular community by offering productive force and services.
The currency is a commons resource used by the whole community.

The use of networked criptocurrencies evidences that flows of money are nothing more than flows of information.
Money transactions are simply communications among the involved parties.
The freedom to use different currencies is directly related to the freedom to communicate, which is popularly referred to as free speech.
The flows of money within a community are communications flows that strengthen the relation between the members of the community and therefore the community as a whole.

The value of the currency increases as the number of users and transactions increases.
Therefore, the members of a community have an incentive to devote some currency resources to attract new members to the community.
It is in their self-interest to make donations that increase the value of the currency.
In this regards, cryptocurrencies are a transition step towards a gift economy.

The gift (or reputation) economy is the kind of economy we find on the Internet, in which the participants offer software, articles, music, pictures and all kind of digital goods for free using permissive licenses such as Creative Commons.
In the information age in which is obvious that money is just information and wealth is measured in knowledge instead of material goods, the progressive evolution towards a reputation economy is a natural step.

\section{Community-Centric Innovation}

Throughout human history, organizational models have changed, transformed and evolved.
As technology evolves, old models are been progressively replaced by others that more efficient in the context of the new technologies.
The invention of information processing and transmission tools favors community-centric models.
Initiatives embracing the community centric model have an advantage compared to those attaching to old hierarchical approaches.

The first example of successes can be found precisely in the fields of information and communication technologies.
The Internet itself is probably the best example of a community-centric endeavour.
There is a large number of people fascinated by and contributing to the Internet.
The decision-making forums are open and anyone is free to contribute expertise.
In a first level the users organize as networks, and then at another level those networks interconnect among them using peering agreements.

Another example of the successful application of the community-centric model is the free libre opens source movement.
Thousands of programmers affiliate to and contribute to this movement working as self-managed community.
The goods offered by this community often outperform those that have been developed using hierarchical models.
By making it possible for anyone to contribute, FLOSS projects have an advantage with respect to those that choose more restrictive models.

The Wikipedia is an Internet-based project that has also adopted the community-centric model.
The users are not mere consumers of information. 
On the contrary, the community of users is the one playing a central role by writing, revising and correcting the encyclopedical articles.
The community-centric model has allowed Wikipedia to grow in terms of number of articles, languages and quality to a point that surpassed encyclopedias that were built using more traditional models.

There are also examples of success of community-centric models in hardware initiatives.
Both Arduino and Raspberry-Pi are open hardware projects that are aligned with community-centric innovation.
They both share all the information regarding the hardware with the community of users that, in turn, use this information to contribute to the project.
